Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-09 ms.tex
Fleshing out heating section

\begin{abstract}
Steady-state one-dimensional hydrodynamic profiles of hot gas in the nuclei of quiescent galaxies are calculated for a range of black hole masses, parameterized gas heating rates, and physically-motivated stellar density profiles.  Mass is supplied to the circumnuclear medium (CNM) by stellar winds, while gas is heated by a combination of stellar winds, supernovae, and black hole feedback.  Analytic estimates are derived, neglecting radiative cooling, for quantities including the stagnation radius where the radial velocity goes to zero (if it exists) and the black hole accretion rate as a function of the heating efficiency, the latter of which is related to the star-formation history of the stellar population.  These results are used to assess over what range of conditions radiative instability will develop, assuming a feedback due to the average star formation history for galaxies of the given black hole mass.  Our main conclusion is that.....  Implications of our results for estimates of the accretion rate using Bondi formula, and for the diversity of CNM environments encountered by the jetted outflows from tidal disruption events, and for the use of an assumed $\dot{M}-M_{\bullet}$ relationship to constrain the occupation fraction of SMBHs in low mass galaxies, are briefly discussed.
\end{abstract}

@@

To understand why most SMBHs appear to be inactive requires understanding their gaseous environments.  Hot gas near the
SMBH sphere of influence (hereafter the `circumnuclear medium', or CNM) controls the central mass accretion rate $\dot{M}$.  The latter in turn affects the luminosity of the accretion flow and the possible impact of its energy and momentum input (`feedback') on larger scales.  The presence of dense gas in the nucleus can also lead to runaway cooling instabilities, resulting in bursty star-formation or episodic AGN activity (reference Ciotti and others).

Knowledge of how $\dot{M}$ varies with the SMBH mass $\Mbh$ and other properties of the
galactic nucleus informs key questions related to co-evolution of SMBHs and their host galaxies with cosmic time.  SMBH growth in the low redshift universe appears to be dominated by lower mass black holes (e.g.~\citealt{Heckman+04}).  This is often
attributed to cosmological `down sizing' (e.g, \citealt{Gallo+08}), even though the physical processes involved may be distinct from those operating at higher SMBH masses and accretion rates in quasars.

The accretion rates of low mass SMBHs, and their resulting nuclear X-ray luminosities $L_X$, also have implications for observational
constraints on their occupation fraction in low mass galaxies and the
possible existence of intermediate mass black holes.  By combining the average relationship between $\Mbh$ and
$L_X$ measured from a sample of high mass galaxies (\citealt{Gallo+08}) with upper limits on $L_X$ in lower mass
galaxies, \citet{MillerGallo+:2014a}  infer tentative evidence for an occupation fraction less than unity in the latter.  This method, however, relies on extrapolating a power-law fit to the $L_X-\Mbh$ distribution to low $\Mbh$ below the detection threshold.

%The conclusions of
%\citet{MillerGallo+:2014a} are thus sensitive to possible changes in
%the $L_X$-$\Mbh$ relationship at small $\Mbh$.
% AG: In our model we obtain a steeper relationship, but we have difficulty matching the observed relationship above the detection threshold

The gas density of the CNM also influences the emission from stellar tidal disruption events (TDEs), such as the high energy
transient {\it Swift} J1644+57 (\citealt{Levan+11}; \citealt{Bloom+11}, \citealt{Burrows+11}; \citealt{Zauderer+11}).
This event was powered by a transient relativistic jet, which produced
synchrotron radio emission as the jet material was decelerated by
shock interaction with the CNM of the previously quiescent SMBH
(\citealt{Giannios&Metzger11}; \citealt{Zauderer+11}).  Detailed
modeling of J1644+57 showed that the CNM density was
significantly lower than that measured surrounding Sgr A$^{\star}$
(\citealt{Metzger+12}; \citealt{Berger+12}).  However, a TDE jet
injected into a denser CNM would be decelerated more rapidly,
producing substantially different radio emission than in J1644+57.  Variations in the properties of the CNM could thus in
principle help explain why most TDEs appear to be radio quiet
(e.g.~\citealt{Bower+13}; \citealt{VanVelzen+13}).

Gas comprising the CNM can in principle originate from a variety of sources: (1) wind mass loss from (predominantly evolved) stars; (2) star-star collisions; and (3) unbound debris from TDEs.  Stellar wind mass loss is probably the dominant source insofar as collisions are relevant only in very dense stellar environments for young stellar populations (\citealt{Rubin&Loeb11}), while \citet{MacLeod+13} find that TDEs contribute subdominantly to the time-averaged accretion rate of quiescent SMBHs.

\citet{Ho:2009a} estimates the accretion rates onto a sample of early-type galaxies both by using X-ray observations to determine the Bondi accretion rate and using estimates for the mass loss rate of evolved stars.  Both methods lead to the conclusion that the available gas reservoir is more than sufficient to power the observed low-luminosity AGN, assuming the standard $\sim$ 10 per cent radiative efficiency for thin disk accretion.  Several independent strains of evidence now suggest that low-luminosity AGN result from accretion proceeding in a radiatively inefficient mode (\citealt{Yuan&Narayan14}), due either to the advection of gravitationally-released energy across the SMBH horizon (e.g.~\citealt{Narayan&Yi95}) or due to a low efficiency for the gas inflowing on large scales ultimately reaching the SMBH (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

%In fact, the temperature profile is assumed to be flat inside of the
%Bondi radius.  In reality, the cusp in the velocity dispersion near
%the SMBH, should cause a cusp in the gas temperature profile (assuming
%the kinetic energy of stellar winds is efficiently thermalized in
%shocks).
%% AG-give sense of scale.

Another approach to determine the SMBH accretion rate, which we adopt in this paper, is to directly calculate the density, velocity and temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected to the nuclear environment via
stellar winds, while energy input results from several sources including stellar winds, supernovae, and AGN feedback (e.g., \citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).  Unlike previous works focusing on modeling individual galaxies, here we systematically analyze the properties of the CNM across a sample of galaxies with a range of SMBH masses, stellar density profiles, and star formation histories (\citealt{WangMerritt:2004a};
hereafter WM04).

Many past studies have focused on relatively massive galaxies.  A key result of this work is the onset of , and the need for AGN feedback.  Here we instead neglect radiative cooling, focusing on time-dependent models.  This simple approach allows us to systematically and to derive simple analytic expressions that prove useful in exploring their sensitivity on.

The one-dimensional steady state flow of gas is usually characterized
by an inflow-outflow structure, with a critical radius known as the
``stagnation radius" $\rs$ where the radial velocity goes to zero.
Mass loss from stars interior to the stagnation radius is accreted,
while that outside $\rs$ is unbound in an outflow from the nucleus.
The stagnation radius, rather than the Bondi radius, is thus
ultimately responsible for setting the SMBH accretion rate, although (when it exists)
$\rs$ can in many cases lie within a factor of few of the
nominal Bondi radius.
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion. 
The main goal of our analysis is to provide a framework for
interpreting the properties of quiescent galactic nuclei given the
comparatively well understood physics of stellar winds and supernova
heating.

@@

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the gas density profile $\rho$ of the CNM for each galaxy model by solving the equations of one-dimensional (spherically symmetric) time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial
      v}{\partial r}\right) =-\frac{\partial p}{\partial r}- \rho\frac{GM_{\rm enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] +\nabla\cdot(\kappa \nabla T)
, 
\label{eq:dsdt}
\end{align}
where $v$, $T$, $p$ and $s$ are the radial velocity, temperature,
pressure and specific entropy of the gas, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p =
\rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$. {\bf BDM: take $\mu = 0.62$ here for fully ionized gas - what did you do?}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{profiles.eps}
  \caption{\label{fig:profiles}Radial profiles of the CNM density
    ({\it top}), temperature ({\it middle}), velocity ({\it bottom}),
    and interior X-ray luminosity ({\it bottom}) calculated for the
    cusp galaxies ($\Gamma=0.8$) within our grid.  Different colors
    denote different values of the effective wind heating rate,
    $\vwO=1200$ km s$^{-1}$ ({\it blue}), $\vwO$=600 km s$^{-1}$ ({\it
      orange}), and 300 km s$^{-1}$ ({\it green}).  Different
    linestyles correspond to different values of $\Mbh$, $\Mbh=10^6
    \Msun$ ({\it dot-dashed}), $\Mbh=10^7 \Msun$ ({\it solid}), and
    $\Mbh=10^8 \Msun$ ({\it dashed}). Thin lines correspond to cusp
    galaxies ($\gamma$=0.8) and thick lines correspond to core
    galaxies ($gamma$=0.1). Squares mark out the location of
    the stagnation radius for each solution.
 }
\end{figure}

@@

Fig.~\ref{fig:cooling} shows $H/C=(q \vw^2+q v^2)/2$ and $\tcool/\tff$
as a function of radius for all of our solutions. {\bf BDM: change
  plots}, with different colors denoting different values of $\vwO$.
For our highest heating rates we observe that cooling is unimportant
for all solutions across all radii--neither the local nor the global
thermal instability would be present. However,  for our smallest value
of $\vwO$=300 km s$^{-1}$ with $\Mbh=10^8 \Msun$, the solution would be
thermally unstable as $H/C<1$ and $\tcool/\tff<10$. Note that the
precise region of the parameter space which would be thermally
unstable would depend on the value of $\eta$--in
Fig.~\pageref{fig:cooling} we have chosen $\eta=0.02$, which would be
appropriate for a stellar population with $\tage\simeq \th$. This
value is on the low end of what we would physically expect for
$\eta$.

@@

Main sequence winds, though contributing a much smaller quantity of
mass, have much higher velocities and hence contribute a
disproportionally large share of the kinetic energy of winds.
\citet{NaimanSoares-Furtado+:2013a} estimate that $v_{\rm w,0} \sim
100-150$ km s$^{-1}$ for main sequence winds assuming a relatively old
stellar population.  On the other hand, for a very young stellar
population of age $t_{\star} \lesssim 10^{7}$ yrs the effective wind
heating from massive stars can be much greater, $\gtrsim 10^{3}$ km
s$^{-1}$.  In what follows below, we take an effective stellar wind
heating rate \be v_{\rm w,0}^{\star} = 150\,{\rm km\,s^{-1}} \ee In
Appendix \ref{app:windheat} we present a toy model for how $v_{\rm
  w,0}$ varies as a function of stellar age ({\bf BDM: Nick can
  create}).

\subsection{Millisecond Pulsars} Energy injection from millisecond pulsars (MSPs) is a potentially important heating source.  If the number of MSPs per unit stellar mass is $n_{\rm msp}$ and each contributes on average a spin-down luminosity $\bar{L}_{\rm sd}$, then the resulting heating per unit volume $e \approx \bar{L}_{\rm sd}n_{\rm
  msp}\epsilon$ results in an effective heating rate (eq.~ [\ref{eq:vw_eff}])
\begin{eqnarray}
  v_{\rm w,0}^{\rm MSP} \approx 30\left(\frac{\epsilon}{0.1}\right)^{1/2}\left(\frac{\bar{L}_{\rm
        sd}}{10^{34}\,{\rm erg\,s^{-1}}}\right)^{1/2}
  \eta_{0.02}^{-1/2}\,{\rm km\,s^{-1}},
  \end{eqnarray}
where $\epsilon$ is the thermalization efficiency of the wind, normalized to a value $\epsilon \lesssim 0.1$ inferred based on modeling the insterstellar mediums of globular clusters \citep{NaimanSoares-Furtado+:2013a}.  The above estimate adopts a pulsar density $n_{\rm msp}\simeq3 \times 10^{-40} $ MSPs g$^{-1}$ motivated by the $\sim 30,000$ estimated MSPs in the Milky Way of stellar mass $6\times 10^{10}M_{\odot}$ (\citealt{Lorimer13}).

@@

The total heating rate $\vwO = \sqrt{v_{\rm w,0}^{\star} + v_{\rm w,0}^{\rm Ia} + v_{\rm w,0}^{\rm p} + v_{\rm w,0}^{\bullet}}$ has contributions from stellar winds, supernovae, pulsars, and SMBH feedback.  Since pulsar heating appears unlikely to contribute appreciably, and since SMBH heating is uncertain, we first focus on heating from stellar winds and supernovae.

Figure~\ref{fig:vweff} shows an estimate of how the combined heating
depends on the age of the stellar population $\tage$ and the SMB mass
$\Mbh$.  For $M_{\bullet} \gtrsim 1.5\times 10^8 (\tau_{\star}/t_{\rm
  h})^{0.5} \Msun$, SNe Ia occur sufficiently frequently within the
stagnation radius (i.e., $r_{\rm Ia} > r_{\rm s}$; eq.~[\ref{eq:rIa}])
to contribute their full heating rate.  $v_{\rm w,0}^{\rm Ia} \sim
250$ km s$^{-1}$.  On the other hand, for lower mass BHs Ia SNe are
infrequent, in which case main sequence stars ($v_{\rm w,0}^{\star}
\simeq 100-150$ km s$^{-1}$) dominate the heating at most times of
interest.\footnote{Although SN Ia will still occur surrounding low
  mass SMBHs, if they are infrequent interior to the radius where the
  accretion rate is being set, then they will not affect the flow near
  the SMBH at most epochs.}  For very young stellar populations
($\tage \lsim 10^7$ years), $\vwO\simeq 1000$ km s$^{-1}$ due to
contributions from line driven winds from high mass main sequence
stars.
{\bf AG: We should mention that regime with stellar wind heating only
  is likely to be thermally unstable.}

  \begin{figure}
    \includegraphics[width=\columnwidth]{vw-contour.eps}
    \caption{\label{fig:vweff} Combined heating rate $\vwO$ of the CNM
      in the parameter space of SMBH mass $\Mbh$ and age of the
      stellar population $t_{\star}$.  We assume that stellar winds
      contribute a constant heating rate $100$ km/s. The contribution
      from SNe Ia is calculated using equation (\ref{eq:vw_sne}), with
      thermalization efficiency $\epsilon_{\rm Ia}=0.4$. The
      contribution from SNe Ia is only added if the $r_s$
      corresponding to the resulting $\vwO$ is outside
      $\rIa$.  {\bf BDM: Correct for dependence of Ia heating on time!
      among  other things}}%describe other cases here.
  \end{figure}

  % AG-I am not comfortable saying type Ia SNe's don't contribute to
  % heating for smaller mass BHs. If the cooling time for the SNe is
  % long then it would be wrong to ignore the SNe contribution. The
  % difference would be that the SNe heating would just be less steady.
  % AG-Thornton and http://arxiv.org/pdf/1402.6695v2.pdf...

>>>>

\begin{abstract}
  Steady-state hydrodynamic profiles of hot gas in  the nuclei of a sample of early-type quiescent galaxies are calculated.  We assume
  that mass is supplied to the circumnuclear medium (CNM) entirely by
  stellar winds according to the radial stellar profiles for each
  galaxy, while the gas is heated by a combination of shocked stellar
  winds and supernovae.  Our analytic and numerical results are used
  to estimate the accretion rates onto the central supermassive black
  holes (SMBH) as a function of the SMBH mass $M_{\bullet}$ and the
  heating efficiency, the latter of which is related schematically to
  the age of the stellar population, the frequency of supernovae, and
  the efficacy of accretion feedback processes.  For a fixed heating
  efficiency, we find that the SMBH accretion Eddington factor
  increases with the SMBH mass, $\dot{M}/\dot{M}_{\rm edd} \propto
  M_{\bullet}^{\alpha}$ with $\alpha \approx 0.5, 1$ for cusp and core
  nuclei, respectively.
  % However, considering the potentially lower heating efficiency of low
  % mass SMBHs (due to a younger stellar population and the rareity of
  % supernovae within the stagnation radius) and their resulting higher
  % susceptibility to cooling instability, we find a flatter
  % relationship $\alpha \approx 0$. 
  Our results are used to check previous estimates of the SMBH
  accretion rate assuming Bondi accretion and extrapolating the X-ray
  measured density and temperature to the unresolved Bondi radius.  We
  also discuss the implications of our results for the diversity of
  CNM environments encountered by the jetted outflows from tidal
  disruption events, and for the use of an assumed
  $\dot{M}-M_{\bullet}$ relationship to constrain the occupation
  fraction of SMBHs in low mass galaxies.
\end{abstract}

@@

In order to understand why most SMBHs appear to be inactive, it is
necessary to understand their gaseous environments.  Hot gas near the
SMBH sphere of influence (the `circumnuclear medium' or CNM) controls
the central mass accretion rate $\dot{M}$.  Better knowledge of how
$\dot{M}$ varies with the SMBH mass $\Mbh$ and other properties of the
galactic nucleus is necessary to understand the co-evolution of SMBHs
and their host galaxies with cosmic time.  SMBH growth in the low
redshift universe, for instance, appears to be dominated by low mass
black holes (e.g.~\citealt{Heckman+04}).  Although this is often
attributed to cosmological ``down sizing" (e.g, \citealt{Gallo+08}),
the physical processes involved may be distinct from those operating
at higher SMBH masses and accretion rates in quasars.

The accretion rates of low mass SMBHs and their resulting nuclear
X-ray luminosities $L_X$ also have implications for observational
constraints on their occupation fraction in low mass galaxies and the
possible existence of intermediate mass black holes.
\citet{MillerGallo+:2014a} combine the relationship between $\Mbh$ and
$L_X$ from a sample of relatively high mass galaxies
(\citealt{Gallo+08}) with upper limits on $L_X$ in several lower mass
galaxies, to infer tentative evidence for an occupation fraction less
than unity.  However, this method relies on extrapolating a power-law
fit to the $L_X-\Mbh$ distribution to low $\Mbh$, below the
instrumental detection threshold.  The conclusions of
\citet{MillerGallo+:2014a} are thus sensitive to possible changes in
the $L_X$-$\Mbh$ relationship at small $\Mbh$.
% AG: In our model we obtain a steeper relationship, but we have difficulty matching the observed relationship above the detection threshold

The density profile of the CNM also influences the emission from
stellar tidal disruption events (TDEs), such as the high energy
transient {\it Swift} J1644+57 (\citealt{Levan+11};
\citealt{Bloom+11}, \citealt{Burrows+11}; \citealt{Zauderer+11}).
This event was powered by a transient relativistic jet, which produced
synchrotron radio emission as the jet material was decelerated by
shock interaction with the CNM of the previously quiescent SMBH
(\citealt{Giannios&Metzger11}; \citealt{Zauderer+11}).  Detailed
modeling of {\it Swift} J1644+57 showed that the CNM density was
significantly lower than that measured surrounding Sgr A$^{\star}$
(\citealt{Metzger+12}; \citealt{Berger+12}).  However, a TDE jet
injected into a denser CNM would be decelerated more rapidly,
producing substantially different radio emission than in {\it Swift}
J1644+57.  Variations in the properties of the CNM could thus in
principle help explain why most TDEs appear to be radio quiet
(e.g.~\citealt{Bower+13}; \citealt{VanVelzen+13}).

Gas that forms the CNM of a SMBH may originate from several sources
(\citealt{Ho:2009a}): (1) wind mass loss from (predominantly evolved)
stars; (2) star-star collisions; (3) the unbound debris from TDEs.
Stellar wind mass loss is probably the dominant source in most cases
insofar as collisions are relevant only in very dense stellar
environments for young stellar populations (\citealt{Rubin&Loeb11}),
while TDEs are estimated to provide a subdominant contribution to the
time-averaged accretion rate of quiescent SMBHs
(\citealt{MacLeod+13}).

\citet{Ho:2009a} calculates the accretion rates onto a sample of
early-type galaxies both by using X-ray observations to determine the
Bondi accretion rate, and by estimating the contributions from the
mass loss of evolved stars.  Both methods lead him to conclude that
the available reservoir of gas is more than sufficient to power
low-luminosity AGN, assuming the standard radiative efficiency $\eta
\sim 0.1$ for thin disk accretion.  Indeed, there is now strong
observational evidence that the low levels of AGN activity in
quiescent galaxies results from accretion proceeding in a radiatively
inefficient mode (\citealt{Yuan&Narayan14}), either due to the
gravitationally-released energy being advected across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to only a fraction of the gas
inflowing on large scales ultimately reaching the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

The accretion rates of SMBHs can in principle be estimated directly by
assuming Bondi accretion and using {\it Chandra} X-ray observations to
estimate the gas temperature and density near the Bondi radius.  This
technique has led to the intriguing result that the Bondi accretion
power correlates with the AGN jet power (e.g.,
\citealt{AllenDunn+:2006a}; \citealt{Russell+13};
\citealt{FujitaKawakatu+:2014a}), the latter estimated from the energy
required to inflate the observed cavities within the X-ray emitting
gas.  It is, however, almost never possible to resolve the Bondi
radius, thus requiring all such studies to extrapolate the observed
gas temperature and density profiles to much smaller radial scales.
%In fact, the temperature profile is assumed to be flat inside of the
%Bondi radius.  In reality, the cusp in the velocity dispersion near
%the SMBH, should cause a cusp in the gas temperature profile (assuming
%the kinetic energy of stellar winds is efficiently thermalized in
%shocks).
%% AG-give sense of scale.

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected to the nuclear environment via
stellar winds, while energy input results from a number of sources,
such as wind shock heating and supernovae (e.g.,
\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, however, which focus on modeling the nuclei of
individual galaxies, here we systematically analyze the properties of
the CNM across a large sample of quiescent galaxies with a range of
SMBH masses and other properties (\citealt{WangMerritt:2004a};
hereafter WM04).  For each galaxy, stellar winds deposit mass and
energy across a range of radii calculated directly from the observed
{\it stellar} density profile.

The one-dimensional steady state flow of gas is usually characterized
by an inflow-outflow structure, with a critical radius known as the
``stagnation radius" $\rs$ where the radial velocity goes to zero.
Mass loss from stars interior to the stagnation radius is accreted,
while that outside $\rs$ is unbound in an outflow from the nucleus.
The stagnation radius, rather than the Bondi radius, is thus
ultimately responsible for setting the SMBH accretion rate, although
$\rs$ can in many cases lie within a factor of few of the
nominal Bondi radius.
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion. 
The main goal of our analysis is to provide a framework for
interpreting the properties of quiescent galactic nuclei given the
comparatively well understood physics of stellar winds and supernova
heating.

@@

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the gas density profile $\rho$ of the CNM for each galaxy model by solving the equations of one-dimensional (spherically symmetric) time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial
      v}{\partial r}\right) =-\frac{\partial p}{\partial r}- \rho\frac{GM_{\rm enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] +\nabla\cdot(\kappa \nabla T)
, 
\label{eq:dsdt}
\end{align}
where $v$, $T$, $p$ and $s$ are the radial velocity, temperature,
pressure and specific entropy of the gas, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p =
\rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{profiles.eps}
  \caption{\label{fig:profiles}Radial profiles of the CNM density
    ({\it top}), temperature ({\it middle}), velocity ({\it bottom}),
    and interior X-ray luminosity ({\it bottom}) calculated for a
    sample of galaxies within our grid.  Different colors
    denote different values of the effective wind heating rate,
    $\vwO=1200$ km s$^{-1}$ ({\it blue}), $\vwO$=600 km s$^{-1}$ ({\it
      orange}), and 300 km s$^{-1}$ ({\it green}).  Different
    linestyles correspond to different values of $\Mbh$, $\Mbh=10^6
    \Msun$ ({\it dot-dashed}), $\Mbh=10^7 \Msun$ ({\it solid}), and
    $\Mbh=10^8 \Msun$ ({\it dashed}). Thin lines correspond to cusp
    galaxies ($\Gamma$=0.8) and thick lines correspond to core
    galaxies ($Gamma$=0.1). Squares mark out the location of
    the stagnation radius for each solution.
 }
\end{figure}

@@

Fig.~\ref{fig:cooling} shows $H/C=\frac{1}{2} (q \vw^2+q
v^2)/\left|\Lambda(T) n^2\right|$ and $\tcool/\tff$
as a function of radius for all of our solutions. {\bf BDM: change
  plots}, with different colors denoting different values of $\vwO$.
For our highest heating rates we observe that cooling is unimportant
for all solutions across all radii--neither the local nor the global
thermal instability would be present. However,  for our smallest value
of $\vwO$=300 km s$^{-1}$ with $\Mbh=10^8 \Msun$, the solution would be
thermally unstable as $H/C<1$ and $\tcool/\tff<10$. Note that the
precise region of the parameter space which would be thermally
unstable would depend on the value of $\eta$--in
Fig.~\pageref{fig:cooling} we have chosen $\eta=0.02$, which would be
appropriate for a stellar population with $\tage\simeq \th$. This
value is on the low end of what we would physically expect for
$\eta$.

@@

Main sequence winds, though contributing a much smaller quantity of
mass, have much higher velocities and hence contribute a
disproportionally large share of the kinetic energy of winds.
\citet{NaimanSoares-Furtado+:2013a} estimate that $v_{\rm w,0} \sim
100-150$ km s$^{-1}$ for main sequence winds assuming a relatively old
stellar population.  On the other hand, for a very young stellar
population of age $t_{\star} \lesssim 10^{7}$ yrs the effective wind
heating from massive stars can be much greater, $\gtrsim 10^{3}$ km
s$^{-1}$.  In Appendix \ref{app:windheat} we present a model for
how $v_{\rm w,0}$ varies as a function of stellar age.

\subsection{Millisecond Pulsars}
 Energy injection from millisecond pulsars (MSPs) is a potentially
important heating source.  If the number of MSPs per unit stellar mass
is $n_{\rm msp}$ and each contributes on average a spin-down
luminosity $\bar{L}_{\rm sd}$, then the resulting heating per unit
volume $e \approx \bar{L}_{\rm sd}n_{\rm msp}\epsilon$ results in an
effective heating rate (eq.~ [\ref{eq:vw_eff}])

\begin{eqnarray} v_{\rm w,0}^{\rm MSP} \approx
30\left(\frac{\epsilon}{0.1}\right)^{1/2}\left(\frac{\bar{L}_{\rm
sd}}{10^{34}\,{\rm erg\,s^{-1}}}\right)^{1/2} \eta_{0.02}^{-1/2}\,{\rm
km\,s^{-1}},
 \label{eq:vmsp}
  \end{eqnarray}

where $\epsilon$ is the thermalization efficiency of
the wind, normalized to a value $\epsilon \lesssim 0.1$ inferred based
on modeling the insterstellar mediums of globular clusters
\citep{NaimanSoares-Furtado+:2013a}.  The above estimate adopts a
pulsar density $n_{\rm msp}\simeq3 \times 10^{-40} $ MSPs g$^{-1}$
motivated by the $\sim 30,000$ estimated MSPs in the Milky Way of
stellar mass $6\times 10^{10}M_{\odot}$ (\citealt{Lorimer13}).

@@

The total heating rate $\vwO = \sqrt{v_{\rm w,0}^{\star \, 2} + v_{\rm
    w,0}^{\rm Ia \, 2} + v_{\rm w,0}^{\rm p \, 2} + v_{\rm
    w,0}^{\bullet \, 2}}$ has
contributions from stellar winds, supernovae, pulsars, and SMBH
feedback. %  Since pulsar heating appears unlikely to contribute
% appreciably, and since SMBH heating is uncertain, we first focus on
% heating from stellar winds and supernovae.

Each of these heating sources will depend on the integrated star
formation history. This in term will will vary systematically with
galaxy mass and thus with $\Mbh$.  
%Qualitatively smaller galaxies willhave recent star formation and
% energetic winds from Wolf-Rayet stars will dominate the mass
% budget--$\vwO\gsim 1000 $ km/s.  On the other hand for larger galaxies
% will have older populations, and heating will be dominated by Ia
% supernovae on large scales and SMBH feedback on small scales.
We calculate the star formation history as a function of $\Mbh$ and
the $\vwO$ for each heating source using the cosmic star formation
histories in \citet{MosterNaab+:2013a}.  For $v_{\rm w,0}^{\rm p}$ we
use equation~\eqref{eq:vmsp}. The star formation history simply enters
in through $\eta$. The effective $\eta$ for a given star formation
history will be the convolution of the star formation history with
$\eta_{\rm imp}\simeq $ {\bf AG: should have expression for $\eta_{\rm
    Imp}$ in appendix. Have to be careful as we also have the
  simplified expression for $\eta_{\rm Imp}$} ($\eta$ in the impulsive
limit). For $v_{\rm w,0}^{\rm Ia}$ we use
equation~\eqref{eq:vIa}. Besides $\eta$ there is additional dependence
on the star formation history from the Ia rate--which is a convolution
of the star formation history with the delay-time distribution from
\citet{MaozMannucci+:2012a}. The calculation of $v_{\rm w,0}^{\star}$
is somewhat involved: we leave the detailed description for
Appendix~\ref{app:windheat}. {\bf AG: still need a description of the
  Compton heating. In general, I struggled with the level of detail
  here. I couldn't think of a natural place to put the details for the
  other heating sources}

$\vwO$ from each heating source (stars, MSPs, SNe Ia, and black hole
feedback) is plotted in as a function of $\Mbh$ in
Figure~\ref{fig:vwSources}. $v_{\rm w,0}^{\rm p}$ is negligible across
all $\Mbh$. For $\Mbh\lsim 10^{7.5} \Msun$, have substantial recent
star formation and $v_{\rm w,0}^{\star}$ is dominant. For $\Mbh\gsim
10^{7.5} \Msun$ , the younger stellar population with their energetic
stellar winds is dead, and Compton heating plays an important role.

For reference we also plot the inferred $\eta$ for each $\Mbh$ in
Figure~\ref{fig:eta}

\begin{figure}
\includegraphics[width=\columnwidth]{vwSources.pdf}
\caption{\label{fig:vwSources} $\vwO$ for stellar winds ({\it solid
    blue}), SNe Ia ({\it dashed red}), MSPs ({\it solid orange}), as a
  function of $\Mbh$. Each heating source varies with the star
  formation history, which is a function of $\Mbh$--see
  appendix~\ref{app:windheat} for the detailed description of how we
  calculate $\vwO$ as a function of $\Mbh$. We also show the $\vwO$,
  resulting from stellar winds and MSPs ({\it dashed black}) and stellar
  winds, MSPs, and black hole feedback ({\it solid black}). The heating
  will have a weak dependence on the stellar density power law slope,
  $\Gamma$, since $v_{\rm w,0}^{\bullet \, 2}$ depends on
  $\Gamma$. The solid black curve corresponds to $\Gamma=0.8$
  and the dash-dotted black curve corresponds to $\Gamma=0.1$.}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{eta.pdf}
\caption{\label{fig:eta} $\eta$ vs. $\Mbh$.}
\end{figure}

For each $\Mbh$ we may check whether the solution for corresponding
$\vwO$ and $\eta$ would be thermally stable, using
equation~\eqref{eq:cooling2} {\bf: AG--double check that this is
  equivalent to what I did.}

The combined heating rate from stellar winds and Compton heating is
insufficient to prevent thermal instability for the high $\Mbh$.  This
is shown in Figure~\ref{fig:cooling2}, which shows the heating over
cooling rate at $\rs$ as a function of $\Mbh$. For $\Mbh\gsim 10^8
\Msun$, the cooling rate will exceed the cooling rate at $\rs$ and
and the gas would be thermally unstable.

\begin{figure}
\includegraphics[width=\columnwidth]{cooling2.pdf}
\caption{\label{fig:cooling2} Ratio of heating rate to cooling rate
  $\rs$ versus $\Mbh$ for core ($\Gamma=0.1$) galaxies ({\it dashed})
  and cusp ($\Gamma=0.8$) galaxies ({\it solid}). Ratio is calculated
  using equation~\eqref{eq:cooling2} and the results for $\vwO$ and
  $\eta$ in Figures~\ref{fig:vwSources} and~\ref{fig:eta}}
\end{figure}

Although, in principle the Ia contribution to the heating is
substantial, in practice Ia's occur too infrequently to be a steady
contribution to the heating rate on small scales (e.g. $r_{\rm Ia} >
r_{\rm s}$ as shown in Figure~\ref{fig:rs_rIa}). Thus, SNe Ia heating
by itself would not be able to prevent cooling instability.

\begin{figure}
\includegraphics[width=\columnwidth]{rs_rIa.pdf}
\caption{\label{fig:rs_rIa} $\rIa/\rs$ vs. $\Mbh$. $\rIa$ is
  calculated using equation~\eqref{eq:rIa} {\bf AG: I am just
    referring to the first part of the equation here--I use the
    appropriate convolution the Ia rate here}. $\rs$ is calculated
  using the equation~\eqref{eq:stag_simple} and the $\vwO$
  corresponding to each $\Mbh$--see Figure~\ref{fig:vwSources}. For
  the blue curve we use the total $\vwO$ from stars, MSP, and
  black hole feedback (e.g. the solid black curve in
  Figure~\ref{fig:vwSources}). For the purple curve we also include
  the $\vwO$ from Ia's (the dashed red curve
  Figure~\ref{fig:vwSources}).}
\end{figure}
